Project 1
========================================================

This is my Project 1 for Reproducible Research.

### Loading and preprocessing the data

Below is how I loaded in the data.
```{r, echo=FALSE}
setwd("C:/Users/harric17/Desktop/R Stuff/coursera/rep_rsrch/Project1")
```

```{r, echo=TRUE}
mydata = read.csv("activity.csv")
dim(mydata)
```


First I found the sum of the steps per day after removing records with step value of NA
```{r, echo=TRUE}
aggdata <- aggregate(x = mydata[c("steps")], FUN = sum,by = list(Group.date = mydata$date),na.rm=TRUE)
```


### What is mean total number of steps taken per day?

Here's a histogram of those total steps per day

```{r fig.width=7, fig.height=6 }
hist(aggdata$steps,main="Total Steps Per Day")
```

Here's the mean and median of those total steps per day
```{r, echo=TRUE}
mean(aggdata$steps)
median(aggdata$steps)
```

### What is the average daily activity pattern?

Next I found the average number of steps per 5 minute interval averaged accross all days after removing records with step value of NA

```{r, echo=TRUE}
aggdata <- aggregate(x = mydata[c("steps")], FUN = mean,by = list(Group.interval = mydata$interval),na.rm=TRUE)
```

Heres a time series plot of those average steps per interval

```{r, echo=TRUE}
plot(aggdata$Group.interval,aggdata$steps,type="l")
```

Heres the 5 minute interval with the highest average steps

```{r, echo=TRUE}
print(aggdata[aggdata$steps==max(aggdata$steps),c(1:2)],row.names=FALSE)
```

### Imputing missing values

Heres the total number of records where steps are NA
```{r, echo=TRUE}
sum(is.na(mydata$steps))
```

Next I imputed missing values with the overall mean of steps

```{r, echo=TRUE}
b = mean(mydata$steps,na.rm=T)
mydata[,"steps"][is.na(mydata$steps)]=b
```

Here I found the sum of the steps per day using the imputed data
```{r, echo=TRUE}
aggdata <- aggregate(x = mydata[c("steps")], FUN = sum,by = list(Group.date = mydata$date),na.rm=TRUE)
```

Here's a histogram of total steps per day using the imputed data
```{r fig.width=7, fig.height=6 }
hist(aggdata$steps,main="Total Steps Per Day")
```

Here's the mean and median of total steps per day using the imputed data. looks like this has a normalizing effect on the distribution.  
it increases the mean and median.
```{r, echo=TRUE}
mean(aggdata$steps)
median(aggdata$steps)
```

### Are there differences in activity patterns between weekdays and weekends?

Next I created a weekend/weekday variable using non-imputed data
```{r, echo=TRUE}
mydata = read.csv("activity.csv")
mydata$day=weekdays(as.Date(mydata$date))
mydata$weekday[mydata$day %in% c("Saturday","Sunday")] = "Weekend"
mydata$weekday[!(mydata$day %in% c("Saturday","Sunday"))] = "Weekday"
table(mydata$weekday,mydata$day)
```

Heres a plot of interval vs total steps by weekday/weekend after removing NA values
```{r, echo=TRUE}
aggdata <- aggregate(x = mydata[c("steps")], FUN = sum,by = list(Group.interval = mydata$interval, weekday=mydata$weekday),na.rm=TRUE)
library(lattice)
xyplot(steps~Group.interval|weekday, data=aggdata, type="l")

```




